### 2010-03-01 tex/ms.tex
groups-simulations: abstract

\documentclass[12pt,preprint]{aastex}
\usepackage{amssymb,amsmath}
%\usepackage{color,hyperref}
%\definecolor{linkcolor}{rgb}{0,0,0.25}
%\hypersetup{
%  colorlinks=true,        % false: boxed links; true: colored links
%  linkcolor=linkcolor,    % color of internal links
%  citecolor=linkcolor,    % color of links to bibliography
%  filecolor=linkcolor,    % color of file links
%  urlcolor=linkcolor      % color of external links
%}
\setlength{\emergencystretch}{2em}%No overflowing references
\newcommand{\ie}{i.e.}
\newcommand{\etal}{et al.}
\newcommand{\eg}{e.g.}
\newcommand{\eqnname}{equation}
\newcommand{\figurenames}{\figurename s}
\newcommand{\sectionname}{$\mathsection$}
\newcommand{\tvector}[1]{\boldsymbol{\vec{#1}}}
\newcommand{\vx}{\tvector{x}}
\newcommand{\vv}{\tvector{v}}

\begin{document}

\title{Probing Galactic structure with moving groups and radial velocity surveys}
\author{Jo~Bovy}%\altaffilmark{1,2}}
\affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Place, New York, NY 10003, USA}
\email{jb2777@nyu.edu}
%\altaffiltext{1}{Center for Cosmology and Particle Physics, Department of Physics,
%  New York University, 4 Washington Place, New York, NY 10003, USA}
%\altaffiltext{2}{Correspondence should be addressed to jo.bovy@nyu.edu~.}

\begin{abstract}
Various dynamical scenarios related to the Milky Way's bar or spiral
structure have been proposed to explain the existence of moving
groups---clumps of co-moving stars---in the Solar neighborhood. These
models are hard to distinguish with observations of the local Galactic
neighborhood. We show that future programs that survey the Milky Way
disk's kinematics on a global scale can distinguish between these
models. By integrating orbits of stars in various steady-state and
transient bar and spiral structure potentials that produce the pattern
of moving groups observed in the Solar neighborhood, we predict the
velocity distribution as a function of location in the Galactic
disk. We identify the locations that distinguish most clearly between
the various scenarios. Future infrared radial velocity surveys can
probe the velocity distribution at these locations; We predict the
distribution of radial velocities that can be directly compared to
observations to establish the origin of the moving groups and to
constrain the structural and dynamical parameters of the
non-axisymetric Milky Way potential.
\end{abstract}

\keywords{
  Galaxy: bulge
  ---  
  Galaxy: disk
  ---
  Galaxy: evolution
  ---
  Galaxy: fundamental parameters
  ---
  Galaxy: kinematics and dynamics
  ---
  Galaxy: structure
}

\section{Introduction}

\begin{thebibliography}{}

\bibitem[{{Binney} \& {Tremaine}(2008)}]{binneytremaine}
{Binney},~J. \& {Tremaine},~S., 2008, {Galactic Dynamics: Second Edition}
  (Princeton University Press)

\bibitem[Bovy, Hogg, \& Roweis(2009)]{Bovy09a} Bovy,~J., Hogg,~D.~W., \& Roweis,~S.~T., 2009,
  \apj, 700, 1794

\end{thebibliography}

\end{document}