### 2010-05-20 tex/hercules.tex
groups-simulations: small tweaks

\documentclass[12pt,preprint]{aastex}
\usepackage{amssymb,amsmath}
%\usepackage{color,hyperref}
%\definecolor{linkcolor}{rgb}{0,0,0.25}
%\hypersetup{
%  colorlinks=true,        % false: boxed links; true: colored links
%  linkcolor=linkcolor,    % color of internal links
%  citecolor=linkcolor,    % color of links to bibliography
%  filecolor=linkcolor,    % color of file links
%  urlcolor=linkcolor      % color of external links
%}
\setlength{\emergencystretch}{2em}%No overflowing references
\newcommand{\ie}{i.e.}
\newcommand{\etal}{et al.}
\newcommand{\eg}{e.g.}
\newcommand{\eqnname}{equation}
\newcommand{\figurenames}{\figurename s}
\newcommand{\sectionname}{$\mathsection$}
\newcommand{\tvector}[1]{\boldsymbol{\vec{#1}}}
\newcommand{\vx}{\tvector{x}}
\newcommand{\vv}{\tvector{v}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\Gaia}{\emph{Gaia}}
\newcommand{\apogee}{APOGEE}
\newcommand{\hipparcos}{\emph{Hipparcos}}
\newcommand{\vR}{\ensuremath{v_R}}
\newcommand{\vphi}{\ensuremath{v_{\phi}}}
\newcommand{\vo}{\ensuremath{v_0}}
\newcommand{\Ro}{\ensuremath{R_0}}
\newcommand{\ro}{\Ro}
\newcommand{\Ab}{\ensuremath{A_b}}
\newcommand{\Rb}{\ensuremath{R_b}}
\newcommand{\Omegab}{\ensuremath{\Omega_b}}
\newcommand{\fdehnen}{\ensuremath{f_{\text{Dehnen}}}}
\newcommand{\sigmaR}{\ensuremath{\sigma_R}}
\newcommand{\rE}{\ensuremath{R_e}}
\newcommand{\Lc}{\ensuremath{L_c}}
\newcommand{\Rs}{\ensuremath{R_s}}
\newcommand{\Rsigma}{\ensuremath{R_{\sigma}}}
\newcommand{\Rolr}{\ensuremath{R_{\text{OLR}}}}

\begin{document}

\title{Tracing the Hercules stream around the Galaxy}
\author{Jo~Bovy}%\altaffilmark{1,2}}
\affil{Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Place, New York, NY 10003, USA}
\email{jb2777@nyu.edu}
%\altaffiltext{1}{Center for Cosmology and Particle Physics, Department of Physics,
%  New York University, 4 Washington Place, New York, NY 10003, USA}
%\altaffiltext{2}{Correspondence should be addressed to jo.bovy@nyu.edu~.}

\begin{abstract}
It has been proposed that the Hercules stream, a group of co-moving
stars in the Solar neighborhood offset from the bulk of the velocity
distribution, is the result of resonant interactions between stars in
the outer disk and the Galactic bar. Only seen so far in the immediate
Solar neighborhood, I predict the distribution of stellar velocities
and the run of the Hercules stream as a function of location in the
Galactic disk in a simple model for the Galaxy and the bar that
produces the observed Hercules stream. I further show that the
Hercules feature is strong enough that it can be unambiguously
detected in the distribution of line-of-sight velocities in selected
directions. I quantify the most promising lines-of-sight for detection
in radial velocities; These are at $l \gtrsim 250^{\circ}$. The
predictions presented here are only weakly affected by distance
uncertainties, assumptions about the distribution function of the
stellar disk, and the details of the Galactic potential including the
effect of spiral structure; \Gaia\ and future spectroscopic surveys of
the Galactic disk will be able to robustly test the origin of the
Hercules stream and constrain the properties of the Galactic bar.
\end{abstract}

\keywords{Galaxy: bulge --- Galaxy: disk --- Galaxy: evolution ---
  Galaxy: fundamental parameters --- Galaxy: kinematics and dynamics
  --- Galaxy: structure }

\section{Introduction}

The velocity distribution of nearby stars contains various large
overdensities of co-moving stars that cannot be explained as the
vestiges of the clumpy nature of star formation
\citep[\eg,][]{Bovy10a}. In particular, the Hercules moving group, a
group of stars significantly offset from the bulk of the observed
velocity distribution \citep[\eg,][see
  \figurename~\ref{fig:obs}]{Dehnen98b,Bovy10a}, displays a wide range
of ages and metallicities
\citep{Blaauw70a,raboud98a,caloi99a,Bensby07a}. A compelling model
that has it originating through resonant interactions of stars in the
outer disk with the bar in the central region of the Galaxy has been
put forward \citep{dehnen00a,fux01a}, but any confirmation of this
picture beyond the locallly observed stellar kinematics remains
lacking. An investigation of the Hercules members' color--magnitude
diagram found no evidence for any significant metallicity anomaly with
respect to other local stars \citep{Bovy10a}, even though such a
difference should be expected as the Hercules stars originate from a
few kiloparsec toward the Galactic center where the average
metallicity is higher than in the Solar neighborhood. The absence of
the predicted metallicity anomaly could potentially be explained
through, \eg, radial mixing \citep{sellwood02a}, but it is clear that
chemodynamical modeling of the Galactic disk is currently not up to
the task of separating out these effects.

In this Letter I propose that a robust and unambiguous test of the
bar-resonance model for the Hercules stream is the distinct pattern it
predicts as we trace the velocity distribution of stars near the Solar
radius around the Galaxy. \Gaia\ is in a good position to have the
final word on this, but I also show that the Hercules feature can be
detected in the line-of-sight velocity distribution in selected
directions on the sky. Clear signatures should exist in regions a few
kiloparsec from the Sun, allowing for a clean verification of the
bar-resonance model for the Hercules stream.

\section{Methodology}\label{sec:method}

To simulate the effect of the bar on stellar orbits near the Solar
circle, we follow the approach of \citet{dehnen00a}. For a fixed
position in the Galaxy, we evaluate the velocity distribution function
at values for the radial---toward the Galactic center---and
tangential---in the direction of Galactic rotation---velocity by
backward-integrating these velocities in the Galactic potential to
obtain the intial orbit before bar formation. We posit that at this
epoch the old stellar disk was in a steady-state that can be described
by a simple distribution function that is a function of the energy and
angular momentum of the orbit alone. It follows from the collisionless
Boltzmann equation that the value of the velocity distribution at the
present position and velocity is equal to that of the steady-state
distribution function before bar formation evaluated at the initial
orbit.

Since the Hercules stream is only apparent in the planar motions in
the disk \citep[\eg,][]{Bovy09a}, we only consider a two-dimensional
model for the Galaxy. We use a simple power-law rotation curve to
model the axisymmetric Galactic potential
\begin{equation}
v_c(R) = \vo(R/\Ro)^\beta \,,
\end{equation}
and for most of the simulations below we specify this model further to
a flat rotation curve ($\beta = 0$). The model for the bar is given by
the following potential
\begin{equation}
\Phi_b(R,\phi) = \Ab(t) \cos [2(\phi - \Omegab t)] \times
\left\{ \begin{array}{ll} -(\Rb/R)^3\,, & \mathrm{for}\ R \geq
  \Rb\,,\\ (R/\Rb)^3-2\,, & \mathrm{for}\ R \leq \Rb\,. \end{array}
\right.
\end{equation}
Here \Omegab\ is the pattern speed of the bar and \Rb\ is the bar
radius, which is fixed to be 80\,percent of the bar's corotation
radius. The bar is grown smoothly during a time $t_1$, always set to
half of the total integration time in what follows, following the
prescription
\begin{equation}
\Ab(t) = \alpha \, \frac{\vo^2}{3}\,\left(\frac{\Ro}{\Rb}\right)^3
\times \left\{ \begin{array}{lll}
\frac{3}{16}\xi^5 - \frac{5}{8} \xi^3 + \frac{15}{16} \xi + \frac{1}{2}\,, & \xi \equiv 2\frac{t}{t_1}-1\,, & t \leq t_1\,,\\
1\,, & & t \geq t_1\,.\end{array} \right. 
\end{equation}
A comparison of the local velocity distribution obtained by using this
simple bar potential with that resulting from the more realistic bar
potential of \citet{gardner10a} shows that this simple model is
sufficient to model the bar's influence on stellar orbits in the outer
disk.

For the distribution function of the stellar disk before bar
formation, we use a Dehnen distribution function \citep{dehnen99b}
given by
\begin{equation}\label{eq:fdehnen}
\fdehnen(E,L) \propto \frac{\Sigma(\rE)}{\sigmaR^2(\rE)} \, \exp\left[ \frac{\Omega(\rE)\left[L-\Lc(E)\right]}{\sigmaR^2(\rE)}\right]\,,
\end{equation}
where \rE, \Lc, and $\Omega(\rE)$ are the radius, angular momentum,
and angular frequency, respectively, of the circular orbit with energy
$E$. Using the procedure given in \sectionname~3.2 of
\citet{dehnen99b}, we choose the $\Sigma(R)$ and $\sigmaR(R)$
functions such that they reproduce a disk with exponential surface
density and velocity dispersion profiles
\begin{equation}
\Sigma(R) = \Sigma_0 \exp\left(-R/\Rs\right)\,, \qquad 
\sigmaR(R) = \sigma_0 \exp\left(-R/\Rsigma\right)\,,
\end{equation}
to an accuracy of a fraction of a percent at all radii. Our fiducial
model for the distribution function has \Rs = \Ro /3 and \Rsigma =
3\Rs, and $\sigma_0$ such that $\sigmaR(\Ro) = 0.2 \vo$.

As shown by \citet{dehnen99c,dehnen00a}, the model described above
with the Sun at \Ro = 0.9\Rolr---where \Rolr\ is the radius at which a
circular orbit is in the outer Lindblad resonance with the bar---at a
current angle of 25$^{\circ}$ with the bar, with a bar strength
$\alpha = 0.01$, and integrating for four bar periods, reproduces the
Hercules feature in the local velocity distribution while agreeing
with other photometric and kinematical observations of the bar
\citep[\eg,][]{binney97a,bissantz02a,cole02a,Shen10a}. In what
follows, unless indicated otherwise, we fix all of the parameters of
the model at these fiducial values. All velocities are always
considered to be with respect to their local standard of rest in the
absence of a bar.

\section{The full planar velocity distribution in the disk}\label{sec:2d}

\citet{dehnen00a} only simulated the velocity distribution at the
position of the Sun for various combinations of the model parameters
defined in the previous Section. Nevertheless, his results on how the
velocity distribution at the Sun changes with the angle between the
Sun-Galactic center line and the bar could also be interpreted as
showing the velocity distribution on the Solar circle for different
Galactocentric azimuths, fixing the bar angle. In this way, these
simulations become predictions for what one might find when one
studies the velocity distribution in regions beyond those probed by
the \hipparcos\ mission.

In \figurename~\ref{fig:rphi2d}, I present the predictions for the
velocity distributions near the Solar circle for the full
360$^{\circ}$ range in Galactocentric azimuth for the fiducial model
from \sectionname~\ref{sec:method}. Only the range $-90^{\circ} \leq
\phi \leq 90^{\circ}$ is shown here, since the barred Galaxy is
symmetric with respect to rotations through 180$^{\circ}$, such that
the predictions for the other side of the Galaxy can be obtained from
those shown by shifting the azimuths by 180$^{\circ}$. The $R = \Ro$
locus is not shown here, but it can be inferred from \figurename~2 in
\citet{dehnen00a}.

It is clear from this \figurename\ that we can trace the Hercules
stream going around the Galaxy and when looking further in toward the
Galactic center or further out. The effect of the bar is also not
confined to a narrow range in radii, rather the effects of the bar and
especially its outer Lindblad resonance are felt through a large range
of radii (0.6 $\leq R/\Ro \leq 1.4$). These effects are strong and
robust predictions of the bar-origin model for the Hercules stream,
especially in regions beyond the Solar radius, where the Hercules
feature can be found significantly removed from the bulk of the
velocity distribution, which might be severely shaped at all locations
by the processes that create the pattern of low-velocity moving groups
in the local velocity distribution (the main mode in
\figurename~\ref{fig:obs}). As I discuss below, \Gaia\ will be able to
verify these predictions.

\section{Line-of-sight velocity distributions}

Rather than wait for a comprehensive survey of the Galactic disk that
measures the full phase-space position of a large number of stars by
measuring proper motions, we can ask whether a targeted survey of
line-of-sight velocities in selected regions of the Galaxy could see a
clear signature of the Hercules feature. That this should be possible
follows from the following argument. The Hercules feature stands out
in the velocity distribution most in its tangential velocity
$\vphi$. On the Solar circle near the Sun we see almost entirely this
Galactocentric tangential velocity in the line-of-sight, so we would
expect to see at least a hint of the Hercules stream in the line-of
sight.

In \figurename~\ref{fig:rphi1d} I show the predicted line-of-sight
velocity distributions in various locations near the Solar circle all
around the Galaxy, obtained by integrating the full two-dimensional
velocity distribution over the tangential velocity. This
\figurename\ is no longer symmetric with respect to shifts of
180$^{\circ}$ degree in azimuth, since the line-of-sight direction
breaks this symmetry. The gray curve is the line-of-sight distribution
one would observe if the targeted stars were in a steady state in an
axisymmetric potential with a Dehnen distribution function
(\eqnname~[\ref{eq:fdehnen}]). Both line-of-sight velocity
distributions are normalized to have equal mass since I assume that
our knowledge of the Galaxy and the tracer population will not be
great enough that we can test predicted differences in the absolute
number of stars at a given location between the axisymmetric and the
bar-resonance predictions. These differences are also largely a
function of the particular form and parameters of the distribution
function that we choose the represent the stellar disk before bar
formation.

We see that the prospects to see the Hercules feature are quite good
on our side of the Galaxy. The Hercules feature is especially strong
in the line-of-sight velocity distribution for Galactic longitudes $l
\geq 250^{\circ}$. This is because of the fortunate coincidence that
for those lines-of-sight we see velocities in the direction along
which the Hercules streams separates most from the bulk of the
velocity distribution (see \figurename~\ref{fig:rphi2d}) and that at
these azimuths the Hercules stream is at the furthest distance from
the rest of the velocity distribution. We see hardly any effect of the
bar on the line-of-sight velocity distributions on the other side of
the Galaxy ($90^{\circ} \leq \phi \leq 270^{\circ}$). This is due to
the fact that on that side we see mostly the Galactocentric radial
velocities, in which the Hercules feature does not separate clearly
from the bulk of the stars. Even though upcoming surveys such as
\apogee\ have the ability to probe the other side of the Galaxy, for
the purposes of detecting the Hercules stream this is uninteresting.

To further analyze the best directions to target to best see the
Hercules feature in the line-of-sight velocities, I have calculated
the line-of-sight velocity distribution on a 100 by 20 grid in
Galactocentric azimuth and radius. As a measure of the difference
between the bar-resonance model and the steady-state, axisymmetric
model, I have calculated the Kullback-Leibler divergence between these
predicted distributions \citep[\eg,][]{mackay}
\begin{equation}
D_{\text{KL}}(P||Q) = \int \dd x\,p(x)\,\ln \frac{p(x)}{q(x)}\,,
\end{equation}
where, in this case, $x$ is the line-of-sight velocity, $p(x)$ is the
predicted distribution for the bar-origin model of the Hercules
stream, and $q(x)$ is the axisymmetric prediction. The
Kullback-Leibler divergence is only one among many possible choices to
quantify the difference between two probability distributions. It can
be interpreted as the expected weight of evidence for the prediction
by the bar-resonance model over the fully axisymmetric prediction that
can be expected from each sampled star. As such, locations with the
maximal Kullback-Leibler divergence are those in which the Hercules
feature stands out most prominently in the line-of-sight.

The result of this analysis is shown in
\figurename~\ref{fig:detect}. I present results both using the full
line-of-sight velocity distributions and only considering the part of
the distribution at $|v_{\text{los}}| \geq 0.15 \vo$. As is clear from
\figurename~\ref{fig:obs}, the local velocity distribution contains
many more features than the Hercules stream. These features are mostly
contained to the main mode of the velocity distribution $|v| \lesssim
0.15 \vo$. Observations indicate that these features are also the
result of dynamical interactions \citep[\eg,][]{Bovy10a,sellwood10a},
albeit with spiral arms rather than the bar. Therefore, we can expect
these or similar features to be present in the velocity distribution
at other locations in the Galaxy as well, and thus, if the effect of
the Hercules stream occurs close to the center of the bulk of the
velocity distribution it might be hard to disentangle from the other
dynamical effects at play. However, these additional dynamical effects
will be much less important for velocities $\gtrsim 0.15
\vo$. \figurename~\ref{fig:detect} shows that there are many
locations, mostly at $l \gtrsim 250^{\circ}$, where the Hercules
stream is clearly offset from the bulk of the velocity distribution
and should be unambiguously detectable.

Finally, in \figurename~\ref{fig:1dvar} I show the influence of
various nuisance and dynamical parameters on the predicted
line-of-sight velocity distribution for the most promising
line-of-sight and distance from the Sun that is more than 3 kpc
away. We see that the effect of distance uncertainties is
negligible. The exact shape and parameters of the stellar distribution
function also only minimally affect the predicted line-of-sight
velocity distribution. Changing the scalelength within its current
uncertainty changes the prediction only in the amplitude of the
effect: The Hercules stream is more prominent for shorter scalelengths
since then there are relatively more stars in the inner Galaxy that
are brought to the Solar circle by the interaction with the bar. Using
a Shu distribution function \citep{shu69a} instead of the Dehnen
distribution function also gives very similar predictions. Changing
the axisymmetric potential by changing the shape of the rotation curve
shifts the peak as expected \citep[see][]{dehnen00a}, but the Hercules
feature nevertheless remains detectable.

\section{Discussion}

\subsection{\Gaia}

The astrometric \Gaia\ mission will measure the parallaxes and proper
motions of up to one billion stars, most of which will be disk
stars. With parallax and proper motion accuracies down to 10
microarcseconds, \Gaia\ will be able to probe the kinematics of the
disk out to several kpc in all directions
\citep{bailerjones08a}. Crucially, it will provide large samples of
stars in many of the promising regions in
\figurename~\ref{fig:rphi2d}, which will be able to verify the
predictions laid out in this paper and will allow the dynamical
properties of the bar to be tightly constrained, something that will
be hard for \Gaia\ to do using direct observations of the bulge due to
the large extinction toward the Galactic center \citep{robin05a}.

\subsection{Radial velocity surveys}

Since the extinction in the optical bandpasses is rather large in the
directions identified as promising in \figurename~\ref{fig:detect}
\citep[$A_V \approx 8$ mag toward $l \approx
  270^{\circ}$;][]{marshall06a}, spectroscopic surveys in the
near-infrared such as \apogee\ are the best bet for seeing the
Hercules feature in the line-of-sight direction. The range of Galactic
longitudes that \apogee\ can target along the disk is $-5^{\circ} \leq
l \leq 250^{\circ}$ (SDSS-III Project Description\footnote{Available
  at \url{http://sdss3.org/collaboration/description.pdf}~.}). From
\figurename~\ref{fig:detect} this limit seems to be right on the edge
of where one might hope to detect the Hercules feature. Since
\apogee\ is expected to start taking data in 2011 and to finish
operations before the first \Gaia\ data release, \apogee\ could trace
the Hercules stream around the Galaxy before \Gaia\ does. Since
\apogee\ will also obtain detailed chemical abundances for its target
stars, this would also allow for a more detailed analysis of the
abundances of stars in the Hercules stream \citep[following,
  \eg,][]{Bensby07a,Bovy10a}, which could shed more light on radial
mixing in the Galactic disk.

\acknowledgements It is a pleasure to thank Mike Blanton and David
Hogg for helpful discussions and assistance.  Financial support for
this project was provided by the National Aeronautics and Space
Administration (grant NNX08AJ48G) and the National Science Foundation
(grant AST-0908357).

\begin{thebibliography}{}

\bibitem[Bailer-Jones(2008)]{bailerjones08a}
  Bailer-Jones, ~C.~A.~L., 2008,
  in IAU Symp.~254, ed.~J.~Andersen, J.~Bland-Hawthorn, and B.~Nordstr\"{o}m, (Dordrecht: Kluwer), 475

\bibitem[Bensby \etal(2007)]{Bensby07a}
  Bensby,~T., \etal, 2007,
  \apjl, 655, L89

\bibitem[Binney, Gerhard, \& Spergel(1997)]{binney97a}
  Binney,~J.~J., Gerhard,~O., \& Spergel,~D.~N., 1997,
  \mnras, 288, 365

\bibitem[Bissantz \& Gerhard(2002)]{bissantz02a}
  Bissantz,~N.~\& Gerhard,~O., 2002,
  \mnras, 330, 591

\bibitem[Blaauw(1970)]{Blaauw70a}
  Blaauw,~A., 1970, in IAU Symp.~38, ed.~W.~Becker \& I.~Kontopoulos (Dordrecht: Reidel), 199

\bibitem[Bovy, Hogg, \& Roweis(2009)]{Bovy09a} Bovy,~J., Hogg,~D.~W., \& Roweis,~S.~T., 2009,
  \apj, 700, 1794

\bibitem[Bovy \& Hogg(2010)]{Bovy10a} Bovy,~J.~\& Hogg,~D.~W., 2010,
  \apj, in press

\bibitem[Caloi \etal(1999)]{caloi99a}
  Caloi,~V., \etal, 1999,
  \aap, 351, 925

\bibitem[Cole \& Weinberg(2002)]{cole02a}
  Cole,~A.~A.~\& Weinberg,~M.~D., 2002,
  \apjl, 574, L43

\bibitem[{{Dehnen}(1998)}]{Dehnen98b}
Dehnen,~W., 1998, \aj, 115, 2384

\bibitem[Dehnen(1999a)]{dehnen99b}
  Dehnen,~W., 1999a, \aj, 118, 1201

\bibitem[Dehnen(1999b)]{dehnen99c}
  Dehnen,~W., 1999b, \apj, 524, L35

\bibitem[Dehnen(2000)]{dehnen00a}
  Dehnen,~W., 2000, \aj, 119, 800

\bibitem[Fux(2001)]{fux01a}
  Fux,~R., 2001,
  \aap, 373, 511

\bibitem[Gardner \& Flynn(2010)]{gardner10a}
  Gardner,~E.~\& Flynn,~C., 2010,
  \mnras, in press

\bibitem[Mackay(2003)]{mackay}
  Mackay,~D.~J.~C., 2003,
  Information theory, inference, and learning algorithms (Cambridge University Press)

\bibitem[Marshall \etal(2006)]{marshall06a}
  Marshall,~D.~J., \etal, 2006,
  \aap, 453, 635

\bibitem[Raboud \etal(1998)]{raboud98a}
  Raboud,~D., \etal, 1998,
  \aap, 335, L61

\bibitem[Robin \etal(2005)]{robin05a}
  Robin,~A.~C., \etal, 2005,
  \aap, 430, 129

\bibitem[Sellwood(2010)]{sellwood10a}
  Sellwood,~J.~A., 2010,
  \mnras, submitted, 
  arXiv:1001.5197v1 [astro-ph.GA]

\bibitem[Sellwood \& Binney(2002)]{sellwood02a}
  Sellwood,~J.~A.~\& Binney,~J.J., 2002,
  \mnras, 336, 785

\bibitem[Shen \etal(2010)]{Shen10a}
  Shen,~J., \etal, 2010, \apj L, submitted

\bibitem[Shu(1969)]{shu69a}
  Shu,~F.~H., 1969,
  \apj, 158, 505

\end{thebibliography}

\clearpage
\begin{figure}
\includegraphics[width=0.5\textwidth]{veldist.eps}
\caption{The velocity distribution of nearby ($\lesssim 100$ pc) stars
  from \citet{Bovy09a}. \vR\ and \vphi\ are the velocities toward the
  Galactic center and in the direction of Galactic rotation,
  respectively, corrected for the Solar motion. The Hercules moving
  group is visible as an overdensity at \vphi $\approx$ -0.2
  $\vo$.}\label{fig:obs}
\end{figure}

\clearpage
\begin{figure}
\includegraphics[width=\textwidth]{rphi2d.ps}
\caption{Two-dimensional velocity distributions along the Solar
  circle. The axes of each subpanel are the same as in
  \figurename~\ref{fig:obs}. The velocity distributions on the other
  side of the Galaxy are obtained by shifting the azimuths by
  180$^{\circ}$.}\label{fig:rphi2d}
\end{figure}

\clearpage
\begin{figure}
\includegraphics[width=\textwidth]{rphi1d.ps}\\
\includegraphics[width=\textwidth]{rphi1d2.ps}
\caption{Line-of-sight velocity distributions along the Solar
  circle. The range in line-of-sight velocities in each subpanel is as
  in \figurename~\ref{fig:1dvar}. The gray curve in each panel is the
  predicted distribution for an axisymmetric, steady-state
  Galaxy.}\label{fig:rphi1d}
\end{figure}

\clearpage
\begin{figure}
\includegraphics[width=0.5\textwidth]{detecta.ps}
\includegraphics[width=0.5\textwidth]{detectb.ps}
\caption{Kullback-Leibler divergence between the predicted
  line-of-sight velocity distributions with and without the
  bar. Darker values have a larger discriminatory potential. The left
  panel uses the full line-of-sight velocity distributions, the right
  panel excludes the main mode of the velocity distribution that is
  likely to be affected by other dynamical and phase-space
  effects. The concentric dashed lines indicate
  constant-distance slices from the Sun, starting at 1 kpc and spaced
  2 kpc apart.  The white dashed lines indicate the boundaries of the
  four quadrants in Galactic longitude.  The white dashed-dotted line
  shows the region of the sky that \apogee\ can see (-5$^{\circ} \leq
  l \leq 250^{\circ}$).}\label{fig:detect}
\end{figure}

\clearpage
\begin{figure}
\includegraphics[width=0.5\textwidth]{distuncertain.ps}
\includegraphics[width=0.5\textwidth]{df.ps}\\
\includegraphics[width=0.5\textwidth]{slope.ps}
\includegraphics[width=0.5\textwidth]{barstrength.ps}
\caption{Variation of the predicted velocity distribution at R = 1.075
  \Ro , $\phi = -21^{\circ}$ (d= 0.4 \Ro, $l$ = 270$^{\circ}$) with
  distance uncertainties, parameters and shape of the distribution
  function, shape of the rotation curve, and bar
  strength.}\label{fig:1dvar}
\end{figure}

\end{document}